---
title: "Model Performance - rpart"
author:
  - name: Hakan Egeli
  - name: Soumyendu Sarkar
date: 'April 24, 2018'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include = FALSE}

# load the project. This will autoload the data and also create training and test datasets
library('ProjectTemplate')
load.project()

```

```{r, echo=FALSE, include=FALSE}

library(rpart)
library(rattle)
library(caret)

require(gridExtra)

```


## Perform some explortory data analysis.

- we looked at the dtree with all variables
- we choos the top variables
- we did iterated the dtree

- we took the same variables and reiterated out lm

- we took out the sales data to ceate an alternate model which does not depend on sales information

- Median Income and Median Income households which we obtained **census data** played a significant role

## Definitely plot (histogram) any response variables.
##Plot at your predictors.

d-tree plots

multiple univariable plots for lm

##Begin the analysis:

##For analysis/explanatory/exploratory projects start by

##For ML projects

##Build a linear model use lm for regression and/or glm( family=binomial. ... ) for classification

##Evaluate your linear model using your model evaluation RMarkdown document, if you are clever you can reuse the template and version it very quickly.

##Knit this to model-performance-linear.[html|pdf|md] where you are keeping your assets

##Make sure you know who to interpret a linear model! Let me know if we have to cover this.

##Build a tree model (rpart)

##plot the tree paying particular attention to the first few nodes, do they make sense?

##Evaluate your tree with you model evaluation RMarkdown document. Knit this to model-performance-rpart.[html|pdf|md]

##Has your model(s) met the model success criteria?


### Credit Limit vs Sales Numbers




# Decision Tree Model

## examine the relevance of all the dependent variables using a Decision Tree

```{r, echo=FALSE}

model <- rpart(CreditLimit ~ ., data = dataset_train)

```

```{r, echo=FALSE}

fancyRpartPlot(model)

```

```{r, echo=FALSE}

summary(model)

```


# what is the best cp to use?

```{r, echo=FALSE}

plotcp(model)

```

```{r, echo=FALSE}

model$cptable[which.min(model$cptable[,"xerror"]),"CP"]

```

# select the top relevant variables and revise our linear model

```{r, echo=FALSE}
model$variable.importance
#importance(model)

```

We have eliminated AvgDaysOfPay, AvgAmtPastDue, MaxDaysPastDue

```{r, echo=FALSE}

model <- rpart(CreditLimit ~ AvgDaysOfPayCategory + DesignBandSalesLast12Mo + SalesLastYr + TotalSalesLast12Mo + AlternativeMetalSalesLast12Mo + MedianIncomeHouseholds + SalesCurrentYr + PlainBandSalesLast12Mo + MedianEarnings + TotalHouseholds + TotalPopulation + NumberOfJewelryStores + RuralUrbanContinuumCode, data = dataset_train, control = rpart.control(cp = 0.0169))

```

```{r, echo=FALSE}

fancyRpartPlot(model)

```

```{r, echo=FALSE}

summary(model)

```

# Backward Elimination; remove the least relevant (highest p) variable and rebuild the model

SalesLastYr            TotalSalesLast12Mo                SalesCurrentYr       DesignBandSalesLast12Mo        PlainBandSalesLast12Mo AlternativeMetalSalesLast12Mo 
                           24                            21                            11                            11                             9                             8 
         AvgDaysOfPayCategory        MedianIncomeHouseholds

```{r, echo=FALSE}

model <- rpart(CreditLimit ~ DesignBandSalesLast12Mo + AvgDaysOfPayCategory + SalesLastYr + TotalSalesLast12Mo + AlternativeMetalSalesLast12Mo + SalesCurrentYr + PlainBandSalesLast12Mo + MedianIncomeHouseholds, data = dataset_train, control = rpart.control(cp = 0.01))

```

```{r, echo=FALSE}

fancyRpartPlot(model)

```

```{r, echo=FALSE}

printcp(model)

```

Relative error : 19489749 * 0.77040

# build a model based on sales related variables

```{r, echo=FALSE}
dataset_train_alt <- dataset_train

dataset_train_alt$CreditLimit <- factor(dataset_train_alt$CreditLimit) 

model <- rpart(CreditLimit ~ DesignBandSalesLast12Mo + AvgDaysOfPayCategory + SalesLastYr + TotalSalesLast12Mo + AlternativeMetalSalesLast12Mo + SalesCurrentYr + PlainBandSalesLast12Mo, data = dataset_train_alt, control = rpart.control(cp = 0.002), method="class")

```

```{r, echo=FALSE}

fancyRpartPlot(model)

```


```{r, echo=FALSE}

predictions <- predict(model, dataset_train_alt, type = "class")

```

# build a confusion matrix

```{r, echo=FALSE}

cm <- confusionMatrix(dataset_train_alt$CreditLimit, predictions)

ggplot(data = data.frame(cm$table), aes(x=Prediction, y=Reference)) +
  geom_tile(aes(fill=Freq)) + 
  scale_fill_gradient(low = "white", high = "steelblue") +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  xlab("Predictions") +
  ylab("Ground Truth (Observations)") +
  geom_text(aes(label = Freq), size = 2) +
  ggtitle("Decison Tree Plot of Confusion Matrix")

```

```{r, echo=FALSE}

print(cm)

```

```{r, echo=FALSE}

printcp(model)

```

Relative Error : 0.67309 * 0.62884



